ST563 Final Project

### 2021-03-11 ST563-FP.Rmd
added code for the first few models added variable selection code

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model

@@

## Variable Selection

@@

@@

@@

@@

## Logistic Regression  
The logistic regression model uses the "logit" function $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$. Using it for binary classification, we round the fitted values either up or down to 1 or 0. The logistic regression function is defined as  
$$
\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
$$

@@

>>>>

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model
library(glmnet) # fitting LASSO

@@

```{r CV}
# Specify the cross-validation method (5-fold CV)
trctrl <- trainControl(method = "cv", number = 5)
```

## Variable Selection
After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model. Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions. In other words, the more a model relies on a variable to make predictions, the more important it is for the model. 
```{r varImp, echo = FALSE}
# Write preliminary model  
model.class.pre <- class ~ .

# Fit preliminary random forest on entire data set
rf.class.pre <- randomForest(model.class.pre, data=wine, ntree=500, mtry=3, importance=TRUE)

# Create data frame of variable importance
df.class.pre <- rf.class.pre$importance %>% as.data.frame()
X <- rownames(df.class.pre)

# Sort the data frame and prep for plot
VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)

# Plot variable importance
vimp.class <- ggplot(VIrf.pre.class.sort[-12,],aes(x=X,y=MeanDecreaseAccuracy))+
              geom_bar(stat = 'identity', position = 'dodge', fill="lightblue")+
              theme(axis.text.x = element_text(angle = 60, hjust = 1))+
              ylab("Mean Decrease in Accuracy") + xlab("Predictor")+
              ggtitle("Variable Importance Plot")
vimp.class
```

```{r}
# Get X matrix
X <- model.matrix(quality ~ ., data = wine)[, -1]

# Fit LASSO model and plot result
lm.lasso <- glmnet::cv.glmnet(X, wine$quality, alpha = 1)
lambda <- lm.lasso$lambda.min
plot(lm.lasso)

# Fit LASSO model with best tune
lasso.best <- glmnet(X, wine$quality, alpha = 1)
predict(lasso.best, s = lambda, type = "coefficients")

# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
smre <- lm.sub %>% summary()

# Plot the Cp, BIC, and R^2 results in one plot
par(mfrow = c(2,2))
plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")

# Best subset selection
best.betas <- coef(lm.sub, id = 6) %>% round(2)
best.betas
```

```{r form}
form.reg <- quality ~ volatile.acidity + chlorides + 
            total.sulfur.dioxide + pH + sulphates + alcohol
form.class <- class ~ alcohol + sulphates + volatile.acidity + density + total.sulfur.dioxide
```

@@

```{r tree, echo=FALSE}
#### REGRESSION ####
# Fit basic regression tree model
tree.reg <- train(form.reg, data = trn, method = "rpart",
                  trControl = trctrl)

# Obtain RMSE and results
tree.reg.pred <- predict(tree.reg, newdata = tst)
tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))

#### HOTEL 2 ####
# Fit classification tree model
tree.class <- train(form.class, data = trn, method = "rpart",
                    trControl = trctrl)

# Obtain confusion matrix and results for H2
tree.class.pred <- predict(tree.class, newdata = tst)
tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])

# Print out regression tree visualization
tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
rpart.plot(tree.reg.out, main="Regression Tree")

# Print out classification tree visualization
tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
rpart.plot(tree.class.out, main="Classification Tree")

# Print out confusion matrix
list("RMSE" = tree.reg.rmse,
     "Class Results" = tree.class.res)
```

@@

```{r rf, echo=FALSE}
#### REGRESSION ####
# Fit the regression random forest model 
rf.reg <- train(form.reg, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

# Obtain RMSE and results
rf.reg.pred <- predict(rf.reg, newdata = tst)
rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)

#### CLASSIFICATION ####
# Fit the classification random forest model
rf.class <- train(form.class, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

# Obtain confusion matrix and results
rf.class.pred <- predict(rf.class, newdata = tst)
rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])

# Print out results in a list
list("RMSE" = rf.reg.rmse,
     "Class Results" = rf.class.res)
```

@@

```{r svm}
#### CLASSIFICATION ####
# Fit support vector machine for H1
svm.class <- train(form.class, data = trn, method = "svmRadial", 
                   preProcess = c("center", "scale"),
                   trControl = trctrl, tuneLength = 10)

# Obtain confusion matrix and results for H1
svm.class.pred <- predict(svm.class, newdata = tst)
svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])

# Print out results in a list
list("Results" = svm.class.res,
     "Confusion Matrix" = svm.class.CM$table)
```

@@

```{r knn}
#### REGRESSION ####
# Fit knn model for H1
knn.reg <- train(form.reg, data = trn, method = "knn",
                 trControl = trctrl, preProcess = c("center", "scale"))

# Obtain RMSE and results
knn.reg.pred <- predict(knn.reg, newdata = tst)
knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)

#### CLASSIFICATION ####
# Fit knn model for H2
knn.class <- train(form.class, data = trn, method = "knn",
                   trControl=trctrl, preProcess = c("center", "scale"))

# Obtain confusion matrix and results for H2
knn.class.pred <- predict(knn.class, newdata = tst)
knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])

# Print out results in a list
list("RMSE" = knn.reg.rmse,
     "Class Results" = knn.class.res)
```

## Logistic Regression  
The logistic regression model uses the "logit" function, $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$. Using it for binary classification, we round the fitted values either up or down to 1 or 0. The logistic regression function is defined as  
$$
\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
$$

@@

```{r logit}
#### CLASSIFICATION ####
# Fit logistic regression model
glm.class <- glm(form.class, data = trn,  family = binomial(link = "logit"))

# Obtain confusion matrix and results for H2
glm.class.pred <- predict(glm.class, newdata = tst, type = "response")
glm.class.pred <- glm.class.pred %>% round() %>% as.factor()
glm.class.CM <- confusionMatrix(glm.class.pred, ifelse(tst$class == "low", 1, 0) %>% as.factor())
glm.class.res <- c(glm.class.CM$overall[1], glm.class.CM$byClass[1:2])

# Print out results in a list
list("Results" = glm.class.res,
     "Confusion Matrix" = glm.class.CM$table)
```
